Re-examining cross-cultural similarity judgments using language statistics

### 2023-05-09 code/4_paper/manuscript/manuscript.Rmd
cogsci ready

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Alternatively, these judgments could be shaped by cross-cultural differences in the *input* to similarity judgment, that is, the statistics of the linguistic and/or physical environment as manifest in everyday experiences.
Perhaps when faced with the triad task, participants from all cultures follow the same process for conceptualizing similarity, but rely on language or culture-specific input to this process.
If we observe a difference in categorization between East Asian and Western participants, it could be that members of both groups use the same procedure (considering similarity that is influenced by both taxonomic and thematic relations), but that the input to this procedure differ between cultures, with East Asian participants exposed to more support for thematic similarity in their experience in comparison to their Western counterparts.
<!-- We might also expect that both conceptualization of similarity and input from experience play a role in driving cross-cultural differences -- East Asian participants may be exposed to more instances of thematic similarity, and prefer to conceptualize similarity as thematic to a greater extent than Westerners. -->

@@

We measured taxonomic versus thematic similarity matching in a forced-choice word triad task in three populations.
Following @Ji2004, we measured preferences in the US and China.
In addition, we collected data in a novel context: Vietnam.
Vietnam is a Southeast Asian country that borders China and has historically been greatly influenced by Chinese culture [@Hui2002].
Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies -- that Eastern and Western cultures have different notions of similarity -- extends beyond mainland China.
In addition to these replication and extension questions, we tested whether fastText vectors from corpora corresponding to each language context (English, Mandarin, and Vietnamese) are good predictors for similarity judgments in each population.

@@

We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants, `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` CN participants, and `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` VN participants, who missed two or more of the 11 attention check questions.

In doing so, we deviated from our preregistered exclusion criterion, which used a more stringent policy, excluding participants who missed any of the 10 attention checks.
This approach would have substantially reduced our sample size, disproportionately affecting the VN sample (resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we performed analyses with both samples and found broadly similar results.
Any differences are noted in the Results & Discussion section.

@@

All stimuli and scripts for the experiment and analysis are available at [blinded for review].
We collected and adapted stimuli from previous studies to create test triads consisting of a cue with one thematic and one taxonomic match option.
For example, "cow," "grass," and "chicken," where "cow" is the cue, "grass" is the thematic match, and "chicken" the taxonomic match.
We included 105 such triads, a superset including triads pulled from supplemental information and in-text examples across the literature, and others that we adapted or created <!--(see SI for the full list of stimuli and sources)-->.
We selected triads on the basis of cultural familiarity within the US, Vietnam, and China.
The triads were translated from English to Vietnamese and Mandarin by fluent bilingual speakers of each language, and back-translated to English by another fluent bilingual who was naive to the original English version.
The back-translations were checked against the original English to identify ambiguities in the translation, which were resolved through discussion and selection of alternative terms when relevant.

@@

```{=tex}
\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue. We elicited similarity match responses by asking "Which thing is most closely related to the bolded [cue] item?" In Mandarin Chinese, we used the phrasing for "most closely related" reported by Ji et al. (2004): "关联". And in Vietnamese, we translated this as "liên quan nhất".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 11 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 11 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}
```
## Corpus model

@@

To give an intuition for our model, consider again the cow: grass/chicken triad.
We retrieved word vectors for "cow" and "grass", and calculated the cosine distance between these vectors.
We did the same for "cow" and "chicken." Our similarity prediction is inversely proportional to the ratio of cosine distance of these pairs.
This is because a larger cosine distance means the word vectors are further apart, and thus the words are less similar.
For example, if the cosine distance of thematic cow-grass is 0.7 and the cosine distance of taxonomic cow-chicken is 0.3 (the more similar of the two), then our model predicts, correspondingly, that 30% of responses to the triad will be grass, and the other 70% chicken.

@@

We observed the strongest preference for thematic matching in China (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), and then the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`; see Figure 1).

@@

For US responding: only the English (EN) corpus was a significant predictor[^2].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^2]: With the preregistered exclusion criterion, only the English
(EN) and Mandarin (ZH) corpus were significant predictors.

For Vietnamese participants' responses, only the Vietnamese (VI) and Mandarin (ZH) corpora were significant predictors[^3].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^3]: With the preregistered exclusion criteria, all three corpora were significant predictors.

@@

We observed some language specificity in this analysis (Figure 2).
The English corpus was the best predictor for US responding, and the Mandarin corpus was the best predictor for Chinese responding.
While this is not the case with the Vietnamese corpus and Vietnamese participants' responding, the Vietnamese corpus was still a significant predictor for this responding.
These results support our hypothesis that specific language statistics can predict cross-cultural variation in similarity judgment.

@@

df.USVNCN_coeffs_cos <- data.frame(country_value, corpus_value, coeff_value_cos, se_value_cos) %>%
  mutate(corpus_value = factor(corpus_value, levels=c("ZH", "VI", "EN")), 
         country_value = factor(country_value, levels=c("CN", "VN", "US")))

plot.coeffs_cos <- ggplot(data = df.USVNCN_coeffs_cos, 
       mapping = aes(x = country_value, y = coeff_value_cos, fill = corpus_value)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin= coeff_value_cos - se_value_cos, ymax = coeff_value_cos + se_value_cos), width=.2,
                 position=position_dodge(.9)) +
  scale_y_reverse() + 
  labs(x = "Country", y = "Fixed effect size", fill = "Corpus") + 
  scale_x_discrete(labels=c("CN" = "China", "VN" = "Vietnam", "US" = "US")) + 
  scale_fill_manual(values=c("#D63230", "#F39237", "#1C77C3"))
```

```{r coeffs_cos, echo=FALSE, warning=FALSE, cache = TRUE, fig.env = "figure", fig.align = "center", set.cap.width=T, num.cols.cap=1, fig.cap="Fixed effect sizes of each corpus lexical statistics (cosine distance proportion) when included as a predictor for China, Vietnam, and US responding, respectively. The English corpus is the best predictor for US response, and the Mandarin corpus is the best predictor for China response."}
#plot.coeffs_freq + 
  plot.coeffs_cos
```

@@

However, in all cultural contexts, adding the other two corpora produced a significantly better fit than the identical model with only the corresponding corpus included as a predictor (US responses: $\chi^2$(`r fit.US_cos_compare["fit.ENVIZH_US_cos", "Df"]`) = `r round(fit.US_cos_compare["fit.ENVIZH_US_cos", "Chisq"], 2)`, p `r ifelse( fit.US_cos_compare["fit.ENVIZH_US_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.US_cos_compare["fit.ENVIZH_US_cos", "Pr(>Chisq)"], 3)), "< .001")`; VN responses: $\chi^2$(`r fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Df"]`) = `r round(fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Chisq"], 2)`, p `r ifelse( fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Pr(>Chisq)"], 3)), "< .001")`; CN responses: $\chi^2$(`r fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Df"]`) = `r round(fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Chisq"], 2)`, p `r ifelse( fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Pr(>Chisq)"], 3)), "< .001")`). This analysis suggests that culture-specific input to similarity judgments (as proxied by language statistics) do not fully explain cross-cultural differences in similarity judgment, whether due to noise or sparsity in the estimates of language statistics or true additional variance picked up by embeddings from other corpora.

@@

In contrast, we found some support for the environmental statistics account: word embeddings based on language-specific corpora were a good predictor for the corresponding country's similarity judgments, even when other corpus statistics were included, and even for filler triads without a thematic/taxonomic structure.
<!-- However, we also found that culture-based conceptualization of similarity uniquely contributes to predicting similarity judgment when included in a model with corpus statistics. --> Overall, our results provide evidence that cross-cultural differences in similarity judgments are related to patterns in linguistic statistics (which also vary across cultures).

@@

There are still many open questions for this account, however. Our operationalization of environmental statistics with word embeddings likely captures both language inputs and culture-specific conceptualization of similarity. Further investigation of the semantic spaces captured by cross-cultural word embedding models is required. Future work should also aim to provide a more specific computational account of how lexical co-occurrence might guide categorization preference beyond the simple proportion-of-similarity model tested here, and investigate other person- or triad-specific factors that might influence how environmental statistics affect similarity judgments.

@@

Blinded for review.

@@

\newpage

>>>>

output: cogsci2016::cogsci_paper
final-submission: \cogscifinalcopy
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Alternatively, these judgments could be shaped by cross-cultural differences in the *input* to similarity judgment, that is, the statistics of the linguistic and/or physical environment as manifest in everyday experiences.
Perhaps when faced with the triad task, participants from all cultures follow the same process for conceptualizing similarity, but rely on language or culture-specific input to this process.
If we observe a difference in categorization between East Asian and Western participants, it could be that members of both groups use the same procedure (considering similarity that is influenced by both taxonomic and thematic relations), but that the input to this procedure differ between cultures, with East Asian participants exposed to more support for thematic similarity in their experience in comparison to their Western counterparts.
***Because this is a question about naturally-occurring differences in real-world reasoning contexts, it is difficult to address experimentally. In a controlled study, we could manipulate participants' input and determine whether differing inputs influence similarity judgments [e.g. @McDonald2001], but this would not address the actual sources of cross-cultural variation in similarity judgments. Here, we draw on real-world data to estimate variation across cultures.***<!-- We might also expect that both conceptualization of similarity and input from experience play a role in driving cross-cultural differences -- East Asian participants may be exposed to more instances of thematic similarity, and prefer to conceptualize similarity as thematic to a greater extent than Westerners. -->

@@

We measured taxonomic versus thematic similarity matching in a forced-choice word triad task in three populations.
Following @Ji2004, we measured preferences in the US and China.
In addition, we collected data in a novel context: Vietnam.
Vietnam is a Southeast Asian country that borders China and has historically been greatly influenced by Chinese culture [@Hui2002]. ***We were unable to find cultural indices specifically on thematic/taxonomic reasoning, but Vietnam shows greater similarity to China than the US across the Hofstede dimensions of cultural comparison [-@hofstede] and is often cited for its similarity to more canonical exemplars of East Asian culture [@Woodside1998; @Hirschman1996; @Whitmore1984], for example, patterning with East Asian countries rather than Southeast Asian countries on a measure of family values [@Minkov2012].*** Therefore, Vietnam serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies -- that Eastern and Western cultures have different notions of similarity -- extends beyond mainland China.
In addition to these replication and extension questions, we tested whether fastText vectors from corpora corresponding to each language context (English, Mandarin, and Vietnamese) are good predictors for similarity judgments in each population.

@@

***We excluded `r df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl` US participants (`r round((df.excl_US$ppts_finished - df.excl_US$ppts_after_att_check_excl) / df.excl_US$ppts_finished * 100, 2)`%), `r df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl` CN participants (`r round((df.excl_CN$ppts_finished - df.excl_CN$ppts_after_att_check_excl) / df.excl_CN$ppts_finished * 100, 2)`%), and `r df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl` VN participants (`r round((df.excl_VN$ppts_finished - df.excl_VN$ppts_after_att_check_excl) / df.excl_VN$ppts_finished * 100, 2)`%[^2]), who missed three or more of the 10 attention check questions.***

[^2]: The high rate of exclusion within our Vietnamese sample may reflect some unfamiliarity with attention checks, which could have produced exploratory responding to these questions.
    However, because exploratory responding to attention checks is indistinguishable from chance responding (from bots or inattentive participants), we did not analyze this data further.

In doing so, we deviated from our preregistered exclusion criterion, which used a more stringent policy, excluding participants who missed any of the 10 attention checks.
This approach would have substantially reduced our sample size, disproportionately affecting the VN sample (resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we performed analyses with both samples and found broadly similar results.
Any differences are noted ***with a footnote*** in the Results & Discussion section.

@@

All stimuli and scripts for the experiment and analysis are available at [blinded for review].
We collected and adapted stimuli from previous studies to create test triads consisting of a cue with one thematic and one taxonomic match option.
For example, "cow," "grass," and "chicken," where "cow" is the cue, "grass" is the thematic match, and "chicken" the taxonomic match.
We included 105 such triads, a superset including triads pulled from supplemental information and in-text examples across the literature, and others that we adapted or created<!--(see SI for the full list of stimuli and sources)-->.
We selected triads on the basis of cultural familiarity within the US, Vietnam, and China.
The triads were translated from English to Vietnamese and Mandarin by fluent bilingual speakers of each language, and back-translated to English by another fluent bilingual who was naive to the original English version.
The back-translations were checked against the original English to identify ambiguities in the translation, which were resolved through discussion and selection of alternative terms when relevant.

@@

```{=tex}
\begin{CJK*}{UTF8}{gbsn}
Each participant completed all 105 triads in blocks of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue. We elicited similarity match responses by asking "Which thing is most closely related to the bolded [cue] item?" In Mandarin Chinese, we used the phrasing for "most closely related" reported by Ji et al. (2004): "关联". And in Vietnamese, we translated this as "liên quan nhất".
The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design.
The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, for example, the cue "bird" with match options "lizard" and "toad." Additionally, we included 10 attention check trials, which were formatted similarly to the test and filler triads but included an instruction instead of a cue item, e.g., "Choose wife" with match options "wife" and "husband." In total, each participant completed 210 similarity judgments (105 test triads and 105 fillers) and 10 attention check questions, with all items presented in randomized orders that varied between subjects.
\end{CJK*}
```
## Corpus model

@@

```{r method, fig.env = "figure", fig.asp = 0.5, fig.align='center', set.cap.width=T, num.cols.cap=1, strip.white=T, fig.cap = "Thematic response prediction based on cosine distance of the triad cow: grass/chicken."}
img <- png::readPNG("figs/model.png")
grid::grid.raster(img)
```

To give an intuition for our model, consider again the cow: grass/chicken triad.
We retrieved word vectors for "cow" and "grass", and calculated the cosine distance between these vectors.
We did the same for "cow" and "chicken." Our similarity prediction is inversely proportional to the ratio of cosine distance of these pairs (Figure 1).
This is because a larger cosine distance means the word vectors are further apart, and thus the words are less similar.
For example, if the cosine distance of thematic cow-grass is 0.7 and the cosine distance of taxonomic cow-chicken is 0.3 (the more similar of the two), then our model predicts, correspondingly, that 30% of responses to the triad will be grass, and the other 70% chicken.

@@

We observed the strongest preference for thematic matching in China (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), and then the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`; see Figure 2).

@@

For US responding: only the English (EN) corpus was a significant predictor[^3].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_US_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_US_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^3]: With the preregistered exclusion criterion, only the English (EN) and Mandarin (ZH) corpus were significant predictors.

For Vietnamese participants' responses, only the Vietnamese (VI) and Mandarin (ZH) corpora were significant predictors[^4].
EN corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_E"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_E", "Pr(>Chisq)"], 3)), "< .001")`. VI corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_V"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_V", "Pr(>Chisq)"], 3)), "< .001")`. ZH corpus: $\beta$ = `r round(fixef(fit.ENVIZH_VN_cos)["theme_cosine_prop_M"], 2)`, $\chi^2$(`r fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Df"]`) = `r round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Chisq"], 2)`, p `r ifelse( fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.ENVIZH_VN_cos.anova["theme_cosine_prop_M", "Pr(>Chisq)"], 3)), "< .001")`.

[^4]: With the preregistered exclusion criteria, all three corpora were significant predictors.

@@

We observed some language specificity in this analysis (Figure 3).
The English corpus was the best predictor for US responding, and the Mandarin corpus was the best predictor for Chinese responding.
While this is not the case with the Vietnamese corpus and Vietnamese participants' responding, the Vietnamese corpus was still a significant predictor for this responding.
These results support our hypothesis that specific language statistics can predict cross-cultural variation in similarity judgment.

@@

is_sig <- c(1, 0, 0, 
            0, 1, 1, 
            0, 0, 1)

df.USVNCN_coeffs_cos <- data.frame(country_value, corpus_value, coeff_value_cos, se_value_cos, is_sig) %>%
  mutate(corpus_value = factor(corpus_value, levels=c("ZH", "VI", "EN")), 
         country_value = factor(country_value, levels=c("CN", "VN", "US")))

plot.coeffs_cos <- ggplot(data = df.USVNCN_coeffs_cos, 
       mapping = aes(x = country_value, y = coeff_value_cos, fill = corpus_value)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin= coeff_value_cos - se_value_cos, ymax = coeff_value_cos + se_value_cos), width=.2,
                 position=position_dodge(.9)) +
  scale_y_reverse() + 
  ylim(0, -9) +
  labs(x = "Country", y = "Fixed effect size", fill = "Corpus") + 
  scale_x_discrete(labels=c("CN" = "China", "VN" = "Vietnam", "US" = "US")) + 
  scale_fill_manual(values=c("#D63230", "#F39237", "#1C77C3")) +
  geom_text(aes(label = ifelse(is_sig, "*", "")), 
            position = position_dodge(width = .9), vjust = -2.25, size = 20 / .pt)
```

```{r coeffs_cos, echo=FALSE, warning=FALSE, cache = FALSE, fig.env = "figure", fig.align = "center", set.cap.width=T, num.cols.cap=1, fig.cap="Fixed effect sizes of each corpus lexical statistics (cosine distance proportion) when included as a predictor for China, Vietnam, and US responding, respectively. Asterisks indicate corpora that are significant predictors. The English corpus is the best predictor for US response, and the Mandarin corpus is the best predictor for China response."}
#plot.coeffs_freq + 
  plot.coeffs_cos
```

@@

However, in all cultural contexts, adding the other two corpora produced a significantly better fit than the identical model with only the corresponding corpus included as a predictor (US responses: $\chi^2$(`r fit.US_cos_compare["fit.ENVIZH_US_cos", "Df"]`) = `r round(fit.US_cos_compare["fit.ENVIZH_US_cos", "Chisq"], 2)`, p `r ifelse( fit.US_cos_compare["fit.ENVIZH_US_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.US_cos_compare["fit.ENVIZH_US_cos", "Pr(>Chisq)"], 3)), "< .001")`; VN responses: $\chi^2$(`r fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Df"]`) = `r round(fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Chisq"], 2)`, p `r ifelse( fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.VN_cos_compare["fit.ENVIZH_VN_cos", "Pr(>Chisq)"], 3)), "< .001")`; CN responses: $\chi^2$(`r fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Df"]`) = `r round(fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Chisq"], 2)`, p `r ifelse( fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Pr(>Chisq)"] > 0.001, paste0("= ", round(fit.CN_cos_compare["fit.ENVIZH_CN_cos", "Pr(>Chisq)"], 3)), "< .001")`). ***This analysis suggests that culture-specific input to similarity judgments (as proxied by language statistics) do not fully explain cross-cultural differences in similarity judgment. This could be due to noise or sparsity in the estimates of language statistics, so that adding additional corpora simply improves fit by expanding the effective size of the model corpus. Alternatively, it could be due to true additional cultural-specific variance picked up by embeddings from other corpora --- this is a question for future work.***

@@

In contrast, we found some support for the environmental statistics account: word embeddings based on language-specific corpora were a good predictor for the corresponding country's similarity judgments, even when other corpus statistics were included, and even for filler triads without a thematic/taxonomic structure.
***This evidence is correlational and cannot evaluate causes of variation in similarity reasoning; indeed, it may be that cultural variation in similarity reasoning shapes the linguistic variation we measure, but we note that this account does not explain the fit to filler triads.***<!-- However, we also found that culture-based conceptualization of similarity uniquely contributes to predicting similarity judgment when included in a model with corpus statistics. --> Overall, our results provide evidence that cross-cultural differences in similarity judgments are related to patterns in linguistic statistics (which also vary across cultures).

@@

There are still many open questions for this account, however.
Our operationalization of environmental statistics with word embeddings likely captures both language inputs and culture-specific conceptualization of similarity.
Further investigation of the semantic spaces captured by cross-cultural word embedding models is required.
Future work should also aim to provide a more specific computational account of how lexical co-occurrence might guide categorization preference beyond the simple proportion-of-similarity model tested here, and investigate other person- or triad-specific factors that might influence how environmental statistics affect similarity judgments.
***This work should also investigate person- and item-specific factors, and characterize responses that are representative of each group compared to the others.***

@@

We are very grateful to Anjie Cao, Mai Nguyen, Victoria Yang, and Van Luong for their help with stimulus design and international participant recruitment. We also thank members of the Language and Cognition Lab for their feedback and helpful comments. This work was funded in part by NSF under grant SBE-2047581, by awards from the McDonnell Foundation and the Center for the Study of Language and Information at Stanford University supporting AC, and by the Vice Provost for Undergraduate Education Small Research Grant at Stanford University supporting KL.

@@
